---
title: "AmazonReviewModeling"
author: "Sofia Caceres"
date: "2026-04-21"
output: html_document
---

```{r setup, include=FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(tidytext)
library(knitr)

results_dir <- "results"
full_models_dir <- file.path(results_dir, "full_models")

rating_counts <- read.csv(file.path(results_dir, "rating_counts.csv"))
word_counts_full <- read.csv(file.path(results_dir, "word_counts_by_group.csv"))
full_model_summary <- read.csv(file.path(full_models_dir, "full_model_data_summary.csv"))
full_model_metrics <- read.csv(file.path(full_models_dir, "full_model_metrics.csv"))
logistic_coefficients <- read.csv(file.path(full_models_dir, "logistic_coefficients.csv"))
random_forest_importance <- read.csv(file.path(full_models_dir, "random_forest_importance.csv"))

total_reviews <- sum(rating_counts$count)

ratings_summary <- rating_counts %>%
  mutate(percent = round(100 * count / total_reviews, 2))

review_word_counts <- word_counts_full %>%
  filter(group %in% c("low", "high")) %>%
  mutate(review_label = if_else(group == "low", "bad", "good")) %>%
  select(review_label, word, n = count)

word_counts <- review_word_counts %>%
  pivot_wider(
    names_from = review_label,
    values_from = n,
    values_fill = 0
  ) %>%
  mutate(
    total = bad + good,
    bad_prop = bad / total,
    good_prop = good / total,
    difference = good_prop - bad_prop
  )
```

# Full Dataset Summary

This report uses the full Amazon review dataset processed with `analyze_reviews.sh`. The original full data file has one truncated final line, so the analysis uses the complete review records.

```{r}
total_reviews
```

```{r}
ratings_summary %>%
  rename(
    Rating = rating,
    Count = count,
    Percent = percent
  ) %>%
  kable()
```

```{r}
ggplot(ratings_summary, aes(x = factor(rating), y = count, fill = factor(rating))) +
  geom_col(show.legend = FALSE) +
  scale_y_continuous(labels = function(x) format(x, big.mark = ",", scientific = FALSE)) +
  labs(
    title = "Distribution of Review Ratings in the Full Amazon Dataset",
    x = "Rating",
    y = "Count"
  )
```

# Word Counts

For the good/bad comparison, 1-star and 2-star reviews are labeled bad, while 4-star and 5-star reviews are labeled good. Three-star reviews are counted in the rating distribution but excluded from the good/bad word comparison.

```{r}
top_words <- review_word_counts %>%
  group_by(review_label) %>%
  slice_max(n, n = 15) %>%
  ungroup()

ggplot(top_words, aes(x = reorder_within(word, n, review_label), y = n, fill = review_label)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ review_label, scales = "free") +
  scale_x_reordered() +
  scale_y_continuous(labels = function(x) format(x, big.mark = ",", scientific = FALSE)) +
  coord_flip() +
  labs(
    title = "Most Common Words in Good vs Bad Amazon Reviews",
    x = "Word",
    y = "Count"
  )
```

```{r}
word_counts %>%
  filter(total >= 1000) %>%
  arrange(difference) %>%
  head(20)

word_counts %>%
  filter(total >= 1000) %>%
  arrange(desc(difference)) %>%
  head(20)
```

```{r}
bad_words <- word_counts %>%
  filter(total >= 1000) %>%
  arrange(difference) %>%
  head(20)

good_words <- word_counts %>%
  filter(total >= 1000) %>%
  arrange(desc(difference)) %>%
  head(20)

bad_words
good_words
```

```{r}
plot_words <- bind_rows(
  bad_words %>% mutate(group = "More common in bad reviews"),
  good_words %>% mutate(group = "More common in good reviews")
)

ggplot(plot_words, aes(x = reorder(word, difference), y = difference, fill = group)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ group, scales = "free_y") +
  coord_flip() +
  labs(
    title = "Words Most Associated with Good vs Bad Reviews",
    x = "Word",
    y = "Good proportion minus bad proportion"
  )
```

# Regression And Random Forest

The logistic regression and random forest models were trained from the terminal with `run_full_models.sh`. The script streams the full Amazon review file, skips the malformed final line, excludes 3-star reviews from the good/bad classification task, and uses the top 100 full-corpus words as binary document features. Identical feature patterns are compressed with sample weights so the models can use all labeled reviews without writing a huge model matrix to disk.

```{r}
full_model_summary %>%
  rename(
    Metric = metric,
    Value = value
  ) %>%
  kable()
```

```{r}
full_model_metrics %>%
  mutate(
    accuracy = round(accuracy, 4)
  ) %>%
  rename(
    Model = model,
    Accuracy = accuracy,
    `Actual good / predicted good` = actual_good_pred_good,
    `Actual good / predicted bad` = actual_good_pred_bad,
    `Actual bad / predicted good` = actual_bad_pred_good,
    `Actual bad / predicted bad` = actual_bad_pred_bad
  ) %>%
  kable()
```

```{r}
model_accuracy <- full_model_metrics %>%
  mutate(
    Model = recode(
      model,
      logistic_regression = "Logistic regression",
      random_forest = "Random forest"
    )
  )

ggplot(model_accuracy, aes(x = Model, y = accuracy, fill = Model)) +
  geom_col(show.legend = FALSE, width = 0.6) +
  geom_text(
    aes(label = paste0(round(accuracy * 100, 1), "%")),
    vjust = -0.4
  ) +
  scale_y_continuous(
    labels = function(x) paste0(round(100 * x), "%"),
    limits = c(0, 1)
  ) +
  labs(
    title = "Model Accuracy on Full Amazon Review Test Set",
    x = "Model",
    y = "Accuracy"
  )
```

```{r}
model_confusion <- bind_rows(
  full_model_metrics %>%
    transmute(model, Actual = "Good", Predicted = "Good", Count = actual_good_pred_good),
  full_model_metrics %>%
    transmute(model, Actual = "Good", Predicted = "Bad", Count = actual_good_pred_bad),
  full_model_metrics %>%
    transmute(model, Actual = "Bad", Predicted = "Good", Count = actual_bad_pred_good),
  full_model_metrics %>%
    transmute(model, Actual = "Bad", Predicted = "Bad", Count = actual_bad_pred_bad)
) %>%
  mutate(
    Model = recode(
      model,
      logistic_regression = "Logistic regression",
      random_forest = "Random forest"
    ),
    Actual = factor(Actual, levels = c("Bad", "Good")),
    Predicted = factor(Predicted, levels = c("Good", "Bad"))
  )

ggplot(model_confusion, aes(x = Predicted, y = Actual, fill = Count)) +
  geom_tile(color = "white") +
  geom_text(aes(label = format(Count, big.mark = ",")), color = "white") +
  facet_wrap(~ Model) +
  coord_equal() +
  scale_fill_gradient(low = "#9ecae1", high = "#08519c") +
  labs(
    title = "Confusion Matrices for Full-Data Models",
    x = "Predicted label",
    y = "Actual label",
    fill = "Count"
  )
```

```{r}
logistic_coefficients %>%
  arrange(desc(coefficient_for_bad_review)) %>%
  head(15) %>%
  rename(
    Word = word,
    `Logistic coefficient for bad review` = coefficient_for_bad_review
  ) %>%
  kable()
```

```{r}
logistic_coefficients %>%
  arrange(coefficient_for_bad_review) %>%
  head(15) %>%
  rename(
    Word = word,
    `Logistic coefficient for bad review` = coefficient_for_bad_review
  ) %>%
  kable()
```

```{r}
logistic_plot_words <- bind_rows(
  logistic_coefficients %>%
    arrange(desc(coefficient_for_bad_review)) %>%
    head(15) %>%
    mutate(group = "More predictive of bad reviews"),
  logistic_coefficients %>%
    arrange(coefficient_for_bad_review) %>%
    head(15) %>%
    mutate(group = "More predictive of good reviews")
)

ggplot(
  logistic_plot_words,
  aes(
    x = reorder_within(word, coefficient_for_bad_review, group),
    y = coefficient_for_bad_review,
    fill = group
  )
) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~ group, scales = "free_y") +
  scale_x_reordered() +
  coord_flip() +
  labs(
    title = "Most Influential Logistic Regression Words",
    x = "Word",
    y = "Coefficient for bad review"
  )
```

```{r}
random_forest_importance %>%
  head(20) %>%
  rename(
    Word = word,
    `Random forest importance` = importance
  ) %>%
  kable()
```

```{r}
random_forest_importance %>%
  head(20) %>%
  ggplot(aes(x = reorder(word, importance), y = importance)) +
  geom_col(fill = "#238b45") +
  coord_flip() +
  labs(
    title = "Top Random Forest Word Importance Scores",
    x = "Word",
    y = "Importance"
  )
```

# Modeling Note

The model results above come from the full terminal run saved in `results/full_models`. One important modeling choice is that the features are binary indicators for whether each top word appears in a review, rather than exact word counts. This keeps the full-dataset model feasible on the available disk space while still using every labeled complete review through sample weights.
